Automatic Metadata Generation for Fish Specimen Image Collections {\footnotesize \thanks{Research supported by NSF OAC \#1940233 and \#1940322.} }

### 2021-04-05 paper/paper_new.tex
Updates from Overleaf

Digitized specimen accessible in open access repositories provide a rich, extensive data source machine-learning and scientific discovery, however, these resources remain largely untapped due image quality and metadata limitations. The overall goal is addressing this need by combining machine-learning and image informatics techniques. The overall goal of our research is to offer a machine-driven [*novel/innovative?? blah...] approach to current manual metadata generation that is prohibitively costly in both terms of labor and time, and to further offer a [*novel] approach to computing higher-level metadata that supports scientific inquiry. Three key objectives are to:
1) Test [*Explore, Demonstrate the us of] Facebook's AI Research's dectectron tool
Specific aims are to use dectectron to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers, as shown in Figure 2.
2) Pursue [*Investigate] image processing at the pixel level.
Pilot testing found the dectectron bounding boxes could be more tightly enclosed the detected objects. We will see if additional masking background as part of the fish, or the
bounding box may clip away part(s) of the fish with pixel analysis methods commonly
found in image informatics. Specific aim [*objectives] of our research are to:
%not sure we have to list all the details, but something like the below, with the 4 aspects of pixel analysis we tested/investigate/experimented with 
1. Identify the appropriate threshold. We will compute optimal scaling between [*or is it Otsu thresholding] to identify an improved threshold value. [Or we will determine is we can computed as the halfway point between the Otsu threshold value and the mean of the background intensities.]
2. Consolidate the foreground
3. Adjusting the bounding box. [We will use computational methods for masking [calcuate a mask?] to determine if bounding box needs to be shrunk or expanded]
% Jane confused here, if this belongs here, or below
4. Developing a fallback approach for when flood fill operations behave unexpectedly the image is too washed out.
3) Compute [Test our ability to calculate] the following higher-level metadata properties: 
1) contrast (intensities of
the foreground and background, 2) centroid and eye center, and if eye is found, the fish 3) side 4) primary axis and clock value, finally the 5) scale and length.
%Following is original Goals/objectives text: The amount of data is immense, with image collections containing 10,000's of entries. These discoveries will only be possible with computational techniques that can process this vast amount of data.

@@

\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's (FAIR) \verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given image. We chose \verb|detectron| for its relative ease of use compared to lower level libraries, and its implementations of powerful architectures developed by FAIR.
For our project, we use it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers,
as shown in Figure \ref{fig:teaser}.

@@

To overcome this under-segmentation, the threshold value should be
either adjusted up or down, depending on if the background is lighter or darker than the fish.
For our current dataset, the background is always lighter (i.e.~closer to 255), so the threshold value needs to be scaled up to include more of the
foreground image.
% Precisely how much to scale this value by is difficult to determine.
For optimal results the scaling should be dependent on the
contrast between the background and foreground,
which can be affected by
% both how washed out the image is and
the level of pigmentation of the fish.
% To take this into account, we compute the mean intensity of the foreground and background pixels, then use these intensity values and the difference between them to scale the threshold value.
% Even with these values in hand it is not entirely clear how to scale the threshold.
We found that an improved threshold value can be computed as the halfway
point between the Otsu threshold value and the
mean of the background intensities.
This adjusted threshold value
% by 50\% of the difference between the mean of the background and the original threshold value, which 
captures most of fish's fins in most cases, without also masking parts of the background.

@@

Since we recalculate the mask of the fish, its centroid must be recalculated
as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}})),
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically, which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance of its blob in \(x\) and \(y\), which yields its principle eigenvector.
The eigenvector can be directly assigned to the property. If an eye is
present, we ensure that \verb|primary_axis| points in the direction of the
eye in relation to the fish's centroid.

@@

\subsection{Ruler Detection}
For all but 2 of the \(7,244\) images \verb|detectron| was able to find the ruler, a nearly perfect correct rate.
In 56 of the images, the ruler itself was found, but the numbers ``2''
and/or the ``3'' on the ruler were not. Therefore a scale calculation could
not be performed, producing a 99.2\% success rate for
the \verb|scale| computation.

@@

\subsection{Side Detection}
\verb|detectron| was unable to find a fish eye
in 246 of the images. Of the remaining \(6,998\) images, the correct side
(\verb|left| or \verb|right|) was detected in all but 6 cases, producing
a 96.5\% correct rate.

% There were six cases where an eye was detected but the \verb|side| value
% was wrong.
For these 6 cases, a spot on the wrong side of the fish was labeled as the
most likely eye within the bounding box of the fish, i.e.~eye detection 
was incorrect. Figure \ref{fig:wrong_eye} presents one such example.
There were an additional 17 images for which the automated process generated a result that did not match the manually created data. For these remaining cases the manual data was incorrect, giving the automated system an error rate \(2.8\) times lower than the human error rate.
This result highlights the additional utility of automated methods to
double-check and verify human-generated metadata.

@@

\begin{table}[H]
    \centering
%    \caption{\verb|foreground.mean| statistics for the three Brightness classes}
    \caption{foreground.mean statistics for the three Brightness classes}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.2 & 13.5 & 1800\\
        Normal & 93.6 & 14.8 & 5186\\
        Bright & 108.4 & 15.2 & 242\\
        \bottomrule
    \end{tabular}
\end{table}
\subsection{Brightness}
Specimen brightness is one of the 22 hand-recorded metadata properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values
correspond to the mean foreground intensity computed by the automated system.
The mean and standard deviation of the mean foreground intensities were
computed for the images in the three human-specified classes.
Table~\ref{tab:bright} contains the resulting values and shows that
automated intensities values provide objective measures that may be used to break images into groups that correlate with human-defined brightness
classifications.
The mean of \verb|foreground.mean| over all 7,244 images is 89.5, with a
standard deviation of 16.6. These numbers suggest, for example, that
\verb|foreground.mean| $\pm$ one-half a standard deviation could define 
\verb|normal| brightness, with values outside this range being classified
as \verb|dark| or \verb|bright|.

@@

\section{Conclusions}
\textbf{Todo}
\subsection{Future Work}
The most pressing next step is to refine the pixel analysis thresholding process so that the entirety of even light colored fish are marked as foreground in the mask. A deficiency of the current process is that it only operates on single channel intensity. Some of the lightest tails appear yellow in hue to the human eye and easily distinguishable, but when compressed to a single intensity value they are almost identical in value to the white background. Considering when the RGB channels of a pixel are not equal in value may be able to improve masking of such features. Another possible approach to solving this is to threshold an mask on subsets of the bounding box, as to ensure that very dark trunk pixels do not affect the thresholding of lighter regions.

Our ultimate goal is to create a generalized process to work on classes of specimen images (specifically all fish images for BGNN), not only the INHS collection. To accomplish this, a much larger training dataset full of more diverse images will be required. Another requirement will be to generalize the ruler reading process beyond the INHS specific reading of digits on the ruler. \textbf{Todo}: Suggest way(s) to do this?
% Todo: Are these worth mentioning?
%Adding things like seahorses and eels.
%Fitting a ``spine'' to specimen.
%\section{Conclusion}
%\textbf{Todo}

>>>>

Digitized specimen accessible in open access repositories provide a rich, extensive data source machine-learning and scientific discovery, however, these resources remain largely untapped due image quality and metadata limitations. The overall goal is addressing this need by combining machine-learning and image informatics techniques. The overall goal of our research is to offer a machine-driven [*novel/innovative?? blah...] approach to current manual metadata generation that is prohibitively costly in both terms of labor and time, and to further offer a [*novel] approach to computing higher-level metadata that supports scientific inquiry. Three key objectives are to:

GOAL 1) Test [*Explore, Demonstrate the us of] Facebook's AI Research's dectectron tool
Specific aims are to use dectectron to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers, as shown in Figure

GOAL 2) Pursue [*Investigate] image processing at the pixel level.
Pilot testing found the dectectron bounding boxes could be more tightly enclosed the detected objects. We will see if additional masking background as part of the fish, or the
bounding box may clip away part(s) of the fish with pixel analysis methods commonly
found in image informatics. Specific aim [*objectives] of our research are to:
%not sure we have to list all the details, but something like the below, with the 4 aspects of pixel analysis we tested/investigate/experimented with

1. Identify the appropriate threshold. We will compute optimal scaling between [*or is it Otsu thresholding] to identify an improved threshold value. [Or we will determine is we can computed as the halfway point between the Otsu threshold value and the mean of the background intensities.]

2. Consolidate the foreground

3. Adjusting the bounding box. [We will use computational methods for masking [calcuate a mask?] to determine if bounding box needs to be shrunk or expanded]
% Jane confused here, if this belongs here, or below

4. Developing a fallback approach for when flood fill operations behave unexpectedly the image is too washed out.

GOAL 3) Compute [*Test our ability to calculate] the following set of higher-level metadata properties: 1) contrast (intensities of the foreground and background, 2) centroid and eye center, and if eye is found, the fish 3) side 4) primary axis and clock value, finally the 5) scale and length.

%Following is original Goals/objectives text: The amount of data is immense, with image collections containing 10,000's of entries. These discoveries will only be possible with computational techniques that can process this vast amount of data.

@@

\subsection{Detectron}
A prerequisite task to performing any advanced metadata property generation
is finding the specimens (and other relevant objects) within the collection
images. Object detection has been a broadly active field of study in recent
years \cite{zou2019object}, and has resulted in a number of well-tested, purpose-built architectures. We elected to use Facebook AI Research's (FAIR) \verb|detectron| tool \cite{wu2019detectron2}, and specifically its implementation of the Mask R-CNN architecture \cite{he2018mask}, for object detection in our project.
\verb|detectron| is built on \verb|pytorch|~\cite{NEURIPS2019_9015} and provides a relatively straightforward method for training on COCO~\cite{DBLP:journals/corr/LinMBHPRDZ14} format datasets. It is able to handle an arbitrary number of object classes, and can classify an arbitrary number of object within a given image. We chose \verb|detectron| for its relative ease of use compared to lower level libraries, and its implementations of powerful architectures developed by FAIR.
For our project, we use it to identify five object classes: fish, fish eyes, rulers, and the numbers 2 and 3 on rulers,
as shown in Figure \ref{fig:teaser}. Objects with a 30\% confidence score or higher are maintained for analysis.

@@

To overcome this under-segmentation, the threshold value should be
either adjusted up or down, depending on if the background is lighter or darker than the fish.
For our current dataset, the background is always lighter (i.e.~closer to 255), so the threshold value needs to be scaled up to include more of the
foreground image.
% Precisely how much to scale this value by is difficult to determine.
For optimal results the scaling should be dependent on the
contrast between the background and foreground,
which can be affected by
% both how washed out the image is and
the level of pigmentation of the fish.
% To take this into account, we compute the mean intensity of the foreground and background pixels, then use these intensity values and the difference between them to scale the threshold value.
% Even with these values in hand it is not entirely clear how to scale the threshold.
We found that an improved threshold value can be computed as the halfway
point between the Otsu threshold value and the
mean of the background intensities.
This adjusted threshold value
% by 50\% of the difference between the mean of the background and the original threshold value, which 
struck a fairly optimal balance between capturing most of fish's fins in most cases, without also masking parts of the background.

@@

Since we recalculate the mask of the fish, its centroid must be recalculated
as well. This can be done via
\begin{equation}
    (\bar{x}, \bar{y}) = (\mathrm{round}(\frac{M_{10}}{M_{00}}), \mathrm{round}(\frac{M_{01}}{M_{00}})),
\end{equation}
where \(M_{00}\) is the pixel area of the fish's blob, \(M_{10}\) is the sum of all the \(x\) values of blob pixels, and \(M_{01}\) is the sum of all the \(y\) values of blob pixels.
\subsubsection{side}
Determining which \verb|side| of the fish is visible is predicated on finding its eye. If an eye is found, the sign of the \(x\) component of the vector from the centroid of the fish to the centroid of the eye specifies which side is up: negative for left and positive for right. This assumes the fish was photographed vertically (i.e.\ dorsal fin on top), which is essentially always the case for all image collections our group has worked on.
\subsubsection{primary\_axis and clock\_value}
The \verb|primary_axis| of a fish can be calculated by taking the covariance of its blob in \(x\) and \(y\), which yields its principle eigenvector.
The eigenvector can be directly assigned to the property. If an eye is
present, we ensure that \verb|primary_axis| points in the direction of the
eye in relation to the fish's centroid.

@@

\subsection{Ruler Detection}
For all but 2 of the \(7,244\) images \verb|detectron| was able to find the ruler, a nearly perfect correct rate.
In 56 of the images, the ruler itself was found, but the numbers ``2''
and/or ``3'' on the ruler were not. Therefore, a scale calculation could
not be performed, producing a 99.2\% success rate for
the \verb|scale| computation.

@@

\subsection{Side Detection}
\verb|detectron| was unable to find a fish eye
in 246 of the images. These eyes were generally extremely dark, small, or looked nothing like those found in the training set. Of the remaining \(6,998\) images, the correct side
(\verb|left| or \verb|right|) was detected in all but 6 cases, producing
a 96.5\% correct rate.

% There were six cases where an eye was detected but the \verb|side| value
% was wrong.
For these 6 cases, a spot on the wrong side of the fish was labeled as the
most likely eye within the bounding box of the fish, i.e.\ eye detection 
was incorrect. Figure \ref{fig:wrong_eye} presents one such example.
There were an additional 17 images for which the automated process generated a result that did not match the manually created data. For these remaining cases the manual data was incorrect, giving the automated system an error rate \(2.8\) times lower than the human error rate.
This result highlights the additional utility of automated methods to
double-check and verify human-generated metadata.

@@

\begin{table}[H]
    \centering
%    \caption{\verb|foreground.mean| statistics for the three Brightness classes}
    \caption{foreground.mean statistics for the three Brightness classes}
    \label{tab:bright}
    \begin{tabular}{cccc}
        \toprule
        \textbf{Class} & \textbf{Mean Intensity} & \textbf{Standard Deviation} & \textbf{Instances}\\
        \midrule
        Dark & 75.2 & 13.5 & 1800\\
        Normal & 93.6 & 14.8 & 5186\\
        Bright & 108.4 & 15.2 & 242\\
        \bottomrule
    \end{tabular}
\end{table}
\subsection{Brightness}
Specimen brightness is one of the 22 hand-recorded metadata properties. It is encoded as \verb|dark|, \verb|normal| or \verb|bright|. These values
correspond to the mean foreground intensity computed by the automated system.
The mean and standard deviation of the foreground intensities were
computed for the images in the three human-specified classes.
Table~\ref{tab:bright} contains the resulting values and shows that
automated intensities values provide objective measures that may be used to break images into groups that correlate with human-defined brightness
classifications.
The mean of \verb|foreground.mean| over all 7,244 images is 89.5, with a
standard deviation of 16.6. These numbers suggest, for example, that
\verb|foreground.mean| $\pm$ one-half a standard deviation could define 
\verb|normal| brightness, with values outside this range being classified
as \verb|dark| or \verb|bright|.

@@

\section{Conclusions}
The BGNN project demonstrates a clear need for improved metadata for specimen image collections. Much effort, time and money have been put into photographing and digitizing physical specimen, but without detailed and complete metadata properties the utility of these repositories for advanced computation analysis and machine learning is limited. Since it is prohibitively time intensive to generate all the pertinent properties by hand, automated techniques are needed to generate this missing metadata at scale.

In this paper we presented such a metadata generation program. It produces 6 of the 22 core BGNN metadata properties~\cite{leipzig2021biodiversity}, as well as image contrast, bounding boxes and scale and length information. Testing this program on \(7,244\) images from the INHS dataset~\cite{INHS}, we see that the vast majority of the resulting metadata is correct within a tolerance of a few percentage points, and in some cases contains fewer mistakes than the manually generated validation metadata. Through further refinement and generalizing beyond only INHS images, we aim to create a tool that can be distributed to specimen image collection curators to correct the metadata sparsity that precipitated this work.
\subsection{Future Work}
The most pressing next step is to refine the pixel analysis thresholding process so that the entirety of even light colored fish are marked as foreground in the mask. A deficiency of the current process is that it only operates on single channel intensity. Some of the lightest tails appear yellow in hue to the human eye and easily distinguishable, but when compressed to a single intensity value they are almost identical in value to the white background. Considering when the RGB channels of a pixel are not equal in value may be able to improve masking of such features. Another possible approach to solving this is to threshold and mask on subsets of the bounding box, as to ensure that very dark trunk pixels do not affect the thresholding of lighter regions.

Our ultimate goal is to create a generalized process to work on classes of specimen images (specifically all fish images for BGNN), not only the INHS collection. To accomplish this, a much larger training dataset full of more diverse images will be required. Another requirement will be to generalize the ruler reading process beyond the INHS specific reading of digits on the ruler, which will likely involve an automated method of reading ruler ticks instead of digits.
% Todo: Are these worth mentioning?
%Adding things like seahorses and eels.
%Fitting a ``spine'' to specimen.
%\section{Conclusion}
%\textbf{Todo}
